D3064R0: OOTA Execution is Provably Vacuous

### 2023-12-12 ootavacuous.tex
Add Memarian citations on C standard
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Because volatile atomics constitute observable
behavior~\cite[\co{intro.abstract}]{ThomasKoeppe2023N4950},
they must be executed in strict accordance with the rules of the abstract
machine.
This means that any volatile atomic operation involving a store
must execute as if all loads whose return values are used to compute
that store's address or value have already returned.
OOTA afficienados will recognize this as a special case of ordering
relaxed loads before relaxed stores, albeit one not requiring
expensive memory-fence instructions on weakly ordered architectures.

@@

We are grateful to David Goldblatt, Jade Alglave, Alan Stern, and Peter
Sewell for their careful review of an early draft of this paper and
to John Wickerson for asking Paul for a rant and taking the proffered
rant seriously.
We also owe David Goldblatt a debt of gratitude for his having asked an
insightful question at the right time, his insights on combinations of
OOTA and UB, and for his ``Deathstation 9000'' demonic CPU.
Alan Stern located many unclear statements and gaps in reasoning.
Martin Uecker contributed valuable insights on backwards-propagating UB.

@@

% One-pager:
% Threats to validity:
% Speed of light might not always be finite.
% Zero-sized atoms might be discovered.
% It might be possible to propagate information at infinite velocity
%	despite the finite speed of light.
% It might be possible to violate causality.
% ML hallucination might convince people that OOTA is OK.

>>>>

The exact definition of a computer language is subject to some debate,
with standards, implementations, and users all having
influence~\cite{KayvanMemarian2016DepthOfC-1,KayvanMemarian2016DepthOfC-2}.
Nevertheless, because volatile atomics constitute observable
behavior~\cite[\co{intro.abstract}]{ThomasKoeppe2023N4950},
they must be executed in strict accordance with the rules of the abstract
machine.
This means that any volatile atomic operation involving a store
must execute as if all loads whose return values are used to compute
that store's address or value have already returned.
OOTA afficienados will recognize this as a special case of ordering
relaxed loads before relaxed stores, albeit one not requiring
expensive memory-fence instructions on weakly ordered architectures.

@@

We are grateful to David Goldblatt, Jade Alglave, Alan Stern, and Peter
Sewell for their careful review of an early draft of this paper and
to John Wickerson for asking Paul for a rant and taking the proffered
rant seriously.
We also owe David Goldblatt a debt of gratitude for his having asked an
insightful question at the right time, his insights on combinations of
OOTA and UB, and for his ``Deathstation 9000'' demonic CPU.
Alan Stern located many unclear statements and gaps in reasoning.
Martin Uecker contributed valuable insights on backwards-propagating UB.
Richard Grisenthwaite patiently explained the architectural constraints
that prevent hardware OOTA.

@@

% One-pager:
% Threats to validity:
% Speed of light might not always be finite.
% Zero-sized atoms might be discovered.
% It might be possible to propagate information at infinite velocity
%	despite the finite speed of light.
% It might be possible to violate causality.
% Hardware and compiler bugs might result in OOTA.
% ML hallucination might convince people that OOTA is OK.
